Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
Spell checking (americanized) removed authors

\author{Dario Cattaruzza, Cristina David, Alessandro Abate, Daniel Kroening}%\email{alessandro.abate@cs.ox.ac.uk}
%\email{dario.cattaruzza@cs.ox.ac.uk}
%\author{Cristina David}\email{cristina.david@cs.ox.ac.uk}
%\email{kroening@cs.ox.ac.uk}

\institute{University of Oxford}

@@

While research on digital control is well
developed~\cite{astrom1997computer}, automated and sound control synthesis
is challenging, particularly when the synthesis objective goes beyond
classical stability.  Recent work~\cite{abate2017automated} described a
method for synthesising safe digital stabilizing controllers for fully
observable LTI plants.  This work differs from the vast majority of existing
approaches in that it explores the continuous space of the system to
synthesise correct by construction minimal digital controllers.  Contrary to
LTL models that require thousands of states and elaborate algorithms to
implement \cite{reissig2017feedback}, these PID-like controllers are represented by a few parameters
and only require a few lines of code, thus reducing the implementation cost
and execution time while remaining resilient to the effects of digitalization.
However,~\cite{abate2017automated}~is restricted to fully observed
systems (i.e., where all the states are known) modelled in discrete time,
and furthemore explores only a limited set of properties that engineers
are interested in. Conversely, 
in most real implementations, the internal state of a plant cannot be
directly measured but rather inferred from its
effect on the output. A plant model that minimises the difference
between its output and the real plant output is called an observer and it is a well known
paradigm in the control literature~\cite{astrom1997computer}.

In this paper we build on the ideas presented in~\cite{abate2017automated}
and develop a new technique that addresses most of these limitations. 
By~combining state-of-the-art synthesis and verification engines we
automatically generate digital observer controllers for a given continuous
plant model that are correct by construction.  %% This approach delivers a high
%% degree of automation, promises to reduce the cost and time of development of
%% digital control dramatically, and requires considerably less expertise than
%% a-posteriori verification.
Specifically, we synthesize stable,
software-implemented embedded controllers along with a model of a physical
plant (observer).  Due to the complexity of such closed-loop systems, we focus
on linear models with known configurations, and perform
parametric synthesis of controllers meeting a number of properties that
define their desired behaviour.

We perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  Our model evaluates the effects of the
quantizers (A/D and D/A converters) and of time discretization, as well as
representation errors introduced by modelling artifacts such as an
observer~\cite{astrom1997computer} in a finite precision domain, while
reasoning about high level properties such as stability, robustness, safety,
and response time, which define the behaviour of the system.  In particular,
safety requirements are frequently overlooked in conventional feedback
control synthesis, but play an important role in systems engineering.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)
\cite{jha-icse10,DBLP:conf/pldi/Solar-LezamaRBE05} approach,
which we enhance with abstraction refinement and optimization
resulting in the novel CEGIS-OR (CEGIS with Optimization and
Refinement).  This design uses an iterative process where each
iteration performs inductive generalisation based on counterexamples
provided by a verification oracle.  Essentially, the inductive
generalisation uses information about a limited number of inputs to
make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems which are initially represented by coarse abstractions
that only get refined on demand.  Consequently, solutions that do not
require a high level of precision are obtained fast.  Since the use
of coarse abstractions may result in low-quality candidate solutions
(i.e. candidates that fail the specification on a very large number of
inputs), after each inductive generalization phase we use optimization
techniques to improve the produced candidates.  In particular, our
optimization procedure aims at reducing the size of the reach space of
the closed loop system, while maintaining other dynamical properties
of interest (safety, stability, and response time).

@@

In summary, this paper makes the following original contributions.
%
\begin{enumerate}
\item We develop a model for inductive synthesis that combines CEGIS with
  abstraction refinement and optimization in order to handle complex
  systems. 
%
\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesise than standard
  full-state feedback controllers (hence the need for abstraction refinement).
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 %
 \item We present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. 
% Sound tools for continuous time reachability are scarce and to the authors'
% knowledge no tool provides a model that combines both. 
\end{enumerate}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear transformations preserve convexity~\cite{boyd2004convex}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer to as
the reach space of the system, generalises the previous notions over an unbounded
time horizon (transitive closure).

@@

A closed-loop digital control interconnection with observer is shown in Figure~\ref{fig:observersystem}. 
The linear feedback controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics, 
and is driven by the observer state $\hat{\vec{x}}$.  
Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the error on the output estimator,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.
Note that the controller and the observer are implemented in a discrete domain 
(quantities within this domain are identified by the index $\langle C \rangle$), 
whereas the plant evolves in a continuous domain. 
We further model the effects of using a discrete controller as the noise signals $\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of the observer state error),
$\nu_2'$ (rounding of the observer state), and $\nu_t$ (output delay of the digital system) 
Lastly, $k_r$ is a reference gain required by the model.  
Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, 
 so that specific closed-loop properties of the overall interconnection are satisfied.

@@

\note{Much like we did in the earlier work, we should emphasise we look at the closed-loop interconnection from the perspective of the digital controller? }

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the plant at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the plant according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

@@

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time.  
A new input value is produced only at every sample time but implemented over the continuous-time plant dynamics. 
%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in \eqref{eq:continuous_tube_param}, 
 and $\mat{A}_d=\mat{A}_{T_s}, \mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with sample time $T_s$, 
 is the state of the model 
 \begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align*}
 \end{theorem}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Modelling errors in the control loop}\label{sec:errors}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

A common occurrence in digital systems is that they do not necessarily satisfy real-time specifications, 
in view of many errors introduced in the dynamics. 
Errors might result from delays due to processing time,
corruptions over communication lines in the control loop, 
and approximations due to digitalisation.

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic disturbance signal ($\nu_t$ in Figure \ref{fig:observersystem}) upper bounded by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.  
\end{theorem}
%
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).

@@

Since we are interested in synthesising controllers with observers, 
in this section we will look at checking stability, safety and the response time in the presence of an observer.
%\footnote{A full description of the theory related to observer design is in Appendix \ref{sec:control_theory}.}

@@

Stability qualitatively denotes the property of a model to evolve within a bounded set around an equilibrium point. 
In this paper, we are interested in asymptotic stability, where the model is further expected to converge to the steady state value (i.e., the equilibrium point).   
%
We specify model stability via Jury's criterion~\cite{fadali}, 
as classical in Control Theory,  %\footnote{More details on Jury's criterion is in Appendix~\ref{sec:closed_stability}.}, 
 which boils down to a requirement expressed as an algebraic formula 
 on the characteristic polynomial of the model.% - a SAT solver will manipulate its coefficients. 
%
In order to facilitate the use of Jury's criterion, we transform the model into
a representation that contains the coefficients of its characteristic
polynomial expressed as sums of controller and plant values. 
The benefit of this representation is that it has no multiplications, which are expensive to handle by SAT solvers (used by the synthesiser). 
%
Such a representation is a model known as 
% is called Controllable and Observable 
``canonical form''~\cite{Astrom08} %\footnote{See parts~\ref{sec:reachable} and~\ref{sec:observable} in the Appendix.} 
and is dynamically equivalent to the given model.  
In this section we explore the implications of using the canonical form, 
alongside the use of FWL representations.

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop. We note that in the case of
stability we only care that their norms are smaller than one, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$ (see Section \ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to look at the colums shared with $\mat{A}_c$.
Thus we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

\begin{theorem}
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p], |\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired portion of the target within which the output must remain after $t_s$.
\end{theorem}

>>>>

\author{%Dario Cattaruzza, Cristina David, Alessandro Abate, Daniel Kroening
}%\email{alessandro.abate@cs.ox.ac.uk}
%\email{dario.cattaruzza@cs.ox.ac.uk}
%\author{Cristina David}\email{cristina.david@cs.ox.ac.uk}
%\email{kroening@cs.ox.ac.uk}

\institute{%University of Oxford
}

@@

While research on digital control is well
developed~\cite{astrom1997computer}, automated and sound control synthesis
is challenging, particularly when the synthesis objective goes beyond
classical stability.  Recent work~\cite{abate2017automated} described a
method for synthesizing safe digital stabilizing controllers for fully
observable LTI plants.  This work differs from the vast majority of existing
approaches in that it explores the continuous space of the system to
synthesize correct by construction minimal digital controllers.  Contrary to
LTL models that require thousands of states and elaborate algorithms to
implement \cite{reissig2017feedback}, these PID-like controllers are represented by a few parameters
and only require a few lines of code, thus reducing the implementation cost
and execution time while remaining resilient to the effects of digitalization.
However,~\cite{abate2017automated}~is restricted to fully observed
systems (i.e., where all the states are known) modeled in discrete time,
and furthermore explores only a limited set of properties that engineers
are interested in. Conversely, 
in most real implementations, the internal state of a plant cannot be
directly measured but rather inferred from its
effect on the output. A plant model that minimizes the difference
between its output and the real plant output is called an observer and it is a well known
paradigm in the control literature~\cite{astrom1997computer}.

In this paper we build on the ideas presented in~\cite{abate2017automated}
and develop a new technique that addresses most of these limitations. 
By~combining state-of-the-art synthesis and verification engines we
automatically generate digital observer controllers for a given continuous
plant model that are correct by construction.  %% This approach delivers a high
%% degree of automation, promises to reduce the cost and time of development of
%% digital control dramatically, and requires considerably less expertise than
%% a-posteriori verification.
Specifically, we synthesize stable,
software-implemented embedded controllers along with a model of a physical
plant (observer).  Due to the complexity of such closed-loop systems, we focus
on linear models with known configurations, and perform
parametric synthesis of controllers meeting a number of properties that
define their desired behavior.

We perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  Our model evaluates the effects of the
quantizers (A/D and D/A converters) and of time discretization, as well as
representation errors introduced by modeling artifacts such as an
observer~\cite{astrom1997computer} in a finite precision domain, while
reasoning about high level properties such as stability, robustness, safety,
and response time, which define the behavior of the system.  In particular,
safety requirements are frequently overlooked in conventional feedback
control synthesis, but play an important role in systems engineering.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)
\cite{jha-icse10,DBLP:conf/pldi/Solar-LezamaRBE05} approach,
which we enhance with abstraction refinement and optimization
resulting in the novel CEGIS-OR (CEGIS with Optimization and
Refinement).  This design uses an iterative process where each
iteration performs inductive generalization based on counterexamples
provided by a verification oracle.  Essentially, the inductive
generalization uses information about a limited number of inputs to
make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems which are initially represented by coarse abstractions
that only get refined on demand.  Consequently, solutions that do not
require a high level of precision are obtained fast.  Since the use
of coarse abstractions may result in low-quality candidate solutions
(i.e. candidates that fail the specification on a very large number of
inputs), after each inductive generalization phase we use optimization
techniques to improve the produced candidates.  In particular, our
optimization procedure aims at reducing the size of the reach space of
the closed loop system, while maintaining other dynamical properties
of interest (safety, stability, and response time).

@@

In summary, this paper makes the following original contributions.
%
\begin{enumerate}
\item We develop a model for inductive synthesis that combines CEGIS with
  abstraction refinement and optimization in order to handle complex
  systems. 
%
\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesize than standard
  full-state feedback controllers (hence the need for abstraction refinement).
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 %
 \item We present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. 
% Sound tools for continuous time reachability are scarce and to the authors'
% knowledge no tool provides a model that combines both. 
\end{enumerate}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear transformations preserve convexity~\cite{boyd2004convex}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer to as
the reach space of the system, generalizes the previous notions over an unbounded
time horizon (transitive closure).

@@

A closed-loop digital control interconnection with observer is shown in Figure~\ref{fig:observersystem}. 
The linear feedback controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics, 
and is driven by the observer state $\hat{\vec{x}}$.  
Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the error on the output estimator,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.
Note that the controller and the observer are implemented in a discrete domain 
(quantities within this domain are identified by the index $\langle C \rangle$), 
whereas the plant evolves in a continuous domain. 
We further model the effects of using a discrete controller as the noise signals $\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of the observer state error),
$\nu_2'$ (rounding of the observer state), and $\nu_t$ (output delay of the digital system) 
Lastly, $k_r$ is a reference gain required by the model.  
Our goal in this work is to synthesize both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, 
 so that specific closed-loop properties of the overall interconnection are satisfied.

@@

%\note{Much like we did in the earlier work, we should emphasise we look at the closed-loop interconnection from the perspective of the digital controller? }

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the plant at integer multiples of the sample time.
 However, in order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the plant according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

@@

The final case we analyze is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time.  
A new input value is produced only at every sample time but implemented over the continuous-time plant dynamics. 
%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in \eqref{eq:continuous_tube_param}, 
 and $\mat{A}_d=\mat{A}_{T_s}, \mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with sample time $T_s$, 
 is the state of the model 
 \begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align*}
 For the proof see Appendix~\ref{sec:proofs}
 \end{theorem}
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Modeling errors in the control loop}\label{sec:errors}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

A common occurrence in digital systems is that they do not necessarily satisfy real-time specifications, 
in view of many errors introduced in the dynamics. 
Errors might result from delays due to processing time,
corruptions over communication lines in the control loop, 
and approximations due to digitalization.

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modeled as a non-deterministic disturbance signal ($\nu_t$ in Figure \ref{fig:observersystem}) upper bounded by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.  
For the proof see Appendix~\ref{sec:proofs}
\end{theorem}
%
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).

@@

Since we are interested in synthesizing controllers with observers, 
in this section we will look at checking stability, safety and the response time in the presence of an observer.
%\footnote{A full description of the theory related to observer design is in Appendix \ref{sec:control_theory}.}

@@

Stability qualitatively denotes the property of a model to evolve within a bounded set around an equilibrium point. 
In this paper, we are interested in asymptotic stability, where the model is further expected to converge to the steady state value (i.e., the equilibrium point).   
%
We specify model stability via Jury's criterion~\cite{fadali}, 
as classical in Control Theory,  %\footnote{More details on Jury's criterion is in Appendix~\ref{sec:closed_stability}.}, 
 which boils down to a requirement expressed as an algebraic formula 
 on the characteristic polynomial of the model.
In order to facilitate the use of Jury's criterion, we transform the model into
a representation that contains the coefficients of its characteristic
polynomial expressed as sums of controller and plant values. 
The benefit of this representation is that it has no multiplications, which are expensive to handle by SAT solvers (used by the synthesizer). 
%
Such a representation is a model known as 
% is called Controllable and Observable 
``canonical form''~\cite{Astrom08} %\footnote{See parts~\ref{sec:reachable} and~\ref{sec:observable} in the Appendix.} 
and is dynamically equivalent to the given model.  
In this section we explore the implications of using the canonical form, 
alongside the use of FWL representations.

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop. We note that in the case of
stability we only care that their norms are smaller than one, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$ (see Section \ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to look at the columns shared with $\mat{A}_c$.
Thus we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

\begin{theorem}
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p], |\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired portion of the target within which the output must remain after $t_s$.  For the proof see Appendix~\ref{sec:proofs}
\end{theorem}

%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
%
We now use this limit to express a constraint in terms of the
coefficients of the characteristic polynomial $P_K'(z)P_L'(z)$.
Let this polynomial be:
%
\begin{equation}
P(z)=\prod_{i=0}^p (z-\lambda_i)=\sum_{j=0}^p c_jz^j.
\end{equation}
%
%We can use Jury's criterion to limit $|\lambda_i| \leq 1$.
Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
to $P'(z)=\prod_{i=0}^p (z-\lambda_i')=\sum_{j=0}^p c_j'z^j$ where $c_j'$ are
unknown.  By replacing $\lambda_i$, we obtain
%
\begin{equation}
P'(z)=\prod_{i=0}^p (z-\lambda_i')=\prod_{i=0}^p \left(z-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^p \frac{c_j}{\overline{\lambda}^{p-j}}z^j, 
\end{equation}
%
where all values are known. Thus, we can ensure that the settling time is smaller than
a certain magnitude by ensuring that 
the eigenvalues are smaller too (which we can do with Jury's criterion).
